---
title: 'ETC5523: Communicating with Data'
subtitle: "Tutorial 7"
author: "Stuart Lee"
date: "Week 7"
output: 
  html_document:
    toc: true
    css: "tutorial-style.css"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

**<span style="color:red">Please do [Preparation](#preparation) *before* the tutorial!</span>**

## `r emo::ji("target")` Objectives

* learn to make a simple R package as a container for data 
* practice writing documentation with `roxygen2`

## `r emo::ji("wrench")` Preparation  

  
1. Install R-packages
```{r}
install.packages(c("devtools", "usethis", "roxygen2"))
```

2. Take a skim of [Chapter 12 - External Data](https://r-pkgs.org/data.html) from the *R Packages* book.

    
##  Exercise 7A

Use the `usethis` package to create a new package called `cwdata`

```{r}
library(usethis)
create_package("cwdata")
```

**Edit the DESCRIPTION file** created and modify the `Authors@R` tag to include 
yourself as the creator and author of the package. 

Add version control to your package and upload your package to your Github.

##  Exercise 7B

Use the `usethis` package to create an R script called "key_crop_yields.R" in the  "data-raw" directory in your packages folder. 

```{r}
use_data_raw(name = "key_crop_yields")
```

The `data-raw` directory is used to hold scripts that generate package data.
The data we will use in this package will be a tidied version of the
Global Crop Yields data you used in tutorial 6. 

A reminder:

> The data below comes from [Our World in Data](https://ourworldindata.org/crop-yields) and sourced from [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-09-01/readme.md). This data contains agricultural yields across crop types and by country from 1960 onwards. 

**Edit the script** to process the data into a long form tidy representation. Here's some code to get you started:

```{r data-raw}
library(readr)
library(janitor)
library(tidyr)
library(magrittr)
url <- 'https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-09-01/key_crop_yields.csv'

key_crop_yields <- read_csv(url) %>% 
  clean_names() %>% 
  pivot_longer(
    cols = -___,
    names_to = "crop", 
    values_to = "tonnes_per_hectare",
    names_pattern = "([^_]+)"
  )

---
```

After you've finished writing your script make sure you've run it.

You should now have a folder called "data" in your package,
with a file called "key_crop_yields.rda".


## Exercise 7C 

We now need to create an R script that lives in the "R" directory
that documents our data object `key_crop_yields`.

```{r}
use_r("key_crop_yields.R")
```

Modify the script and **document the data.**

 - Describe the number of rows and columns 
 - Type of object `key_crop_yields` is. 
 - What are the types of the columns?
 - What do the columns mean?
 - Where did you find it? 

It is possible to use ordinary markdown syntax in documenation with `roxygen2` by running (in fact this is set by default if you used `usethis::create_package()`) 

```{r}
use_roxygen_md()
```

If you get stuck, take a look at [section 12.1.1 of R packages](https://r-pkgs.org/data.html) for guidance.

Once you've finished modifying run

```{r}
devtools::document()
```

at the console to build your package documentation.

## Exercise 7D

We now have a minimally working package that contains some data. 
Restart your R session and then run 

```{r}
devtools::load_all()
```

Now the `key_crop_yields` data is your available to your R session.

**Edit your DESCRIPTION file** to have a more appropriate title and description and version number. 

After you've finished editing run:
```{r}
devtools::check()
```

**Do you get any errors or warnings?**

**Is there any other way you can use Rstudio to perform a check?**

Once you have passed the check, install the package locally:

```{r}
devtools::install()
```

Restart your R session and try loading your package. Check out the help page for the data.

```{r}
library(cwdata)
?key_crop_yields
```

Remember to track and push the changes you've made to package to your github.

# Exercise 7E (Bonus)

Practice writing and documenting some R functions.
